### 2024-06-06 Scripts/Scripts-collection.Rmd
Update Scripts-collection.Rmd

>>>>

```{r Fig s6a}

df_snps_p_corynebacterium  <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\corynebacterium-snps_distribution.csv)")
df_snps_p_corynebacterium$position[df_snps_p_corynebacterium$types =='synonymous_variant'] <- df_snps_p_corynebacterium$position[df_snps_p_corynebacterium$types =='synonymous_variant'] - 5000

p_snps_distribution_corynebacterium <- ggplot(df_snps_p_corynebacterium) + geom_segment(aes(x=position,y=0,xend=position,yend=snps_counts,color=types),data=df_snps_p_corynebacterium,size=.3)  +
  scale_color_manual(values=c('nonsynonymous_variant'='red','synonymous_variant'='blue'),
                     name='',
                     labels = c('Nonsynonymous','Synonymous')) + 
  scale_x_continuous(labels= scales::comma)  +
  guides(color=guide_legend(nrow = 2)) + ggtitle('Corynebacterium striatum')

p_snps_distribution_corynebacterium <- ggpar(p_snps_distribution_corynebacterium,
                                             xlab = "Genomic position",
                                             ylab = "SNP number",
                                             # legend.title = "Groups",
                                             font.xtickslab=c(18),
                                             font.ytickslab=c(18),
                                             font.x = c(18,"bold"),
                                             font.y = c(18,"bold"),
                                             font.legend = c(18,'bold')) + theme_pubr(base_size = 20,legend = "none",margin = FALSE,border = T) + 
  ggeasy::easy_center_title() + theme(plot.title = element_text(face = "italic"))

p_snps_distribution_corynebacterium

###-------------------------------------------------------------------------------------------------------------------------------------------------------------------

###snps distribution for Pseudomonas aeruginosa

df_snps_p_pseudomonas  <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\pseudomonas-snps_distribution.csv)")
df_snps_p_pseudomonas$position[df_snps_p_pseudomonas$types =='synonymous_variant'] <- df_snps_p_pseudomonas$position[df_snps_p_pseudomonas$types =='synonymous_variant'] - 5000

p_snps_distribution_pseudomonas <- ggplot(df_snps_p_pseudomonas) + geom_segment(aes(x=position,y=0,xend=position,yend=snps_counts,color=types),data=df_snps_p_pseudomonas,size=.3)  +
  scale_color_manual(values=c('nonsynonymous_variant'='red','synonymous_variant'='blue'),
                     name='',
                     labels = c('Nonsynonymous','Synonymous')) + 
  scale_x_continuous(labels= scales::comma)  +
  guides(color=guide_legend(nrow = 2)) + ggtitle('Pseudomonas aeruginosa')

p_snps_distribution_pseudomonas <- ggpar(p_snps_distribution_pseudomonas,
                                         xlab = "Genomic position",
                                         ylab = "SNP number",
                                         # legend.title = "Groups",
                                         font.xtickslab=c(18),
                                         font.ytickslab=c(18),
                                         font.x = c(18,"bold"),
                                         font.y = c(18,"bold"),
                                         font.legend = c(18,'bold')) + theme_pubr(base_size = 20,legend = "none",margin = FALSE,border = T) + 
  ggeasy::easy_center_title() + theme(plot.title = element_text(face = "italic"))

p_snps_distribution_pseudomonas

###-------------------------------------------------------------------------------------------------------------------------------------------------------------------

###snps distribution for stenotrophomonas aeruginosa
df_snps_p_stenotrophomonas  <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\stenotrophomonas-snps_distribution.csv)")
df_snps_p_stenotrophomonas$position[df_snps_p_stenotrophomonas$types =='synonymous_variant'] <- df_snps_p_stenotrophomonas$position[df_snps_p_stenotrophomonas$types =='synonymous_variant'] - 5000

p_snps_distribution_stenotrophomonas <- ggplot(df_snps_p_stenotrophomonas) + geom_segment(aes(x=position,y=0,xend=position,yend=snps_counts,color=types),data=df_snps_p_stenotrophomonas,size=.3)  +
  scale_color_manual(values=c('nonsynonymous_variant'='red','synonymous_variant'='blue'),
                     name='',
                     labels = c('Nonsynonymous','Synonymous')) + 
  scale_x_continuous(labels= scales::comma)  +
  guides(color=guide_legend(nrow = 2)) + ggtitle('Stenotrophomonas maltophilia')

p_snps_distribution_stenotrophomonas <- ggpar(p_snps_distribution_stenotrophomonas,
                                              xlab = "Genomic position",
                                              ylab = "SNP number",
                                              # legend.title = "Groups",
                                              font.xtickslab=c(18),
                                              font.ytickslab=c(18),
                                              font.x = c(18,"bold"),
                                              font.y = c(18,"bold"),
                                              font.legend = c(18,'bold')) + theme_pubr(base_size = 20,legend = "none",margin = FALSE,border = T) + 
  ggeasy::easy_center_title() + theme(plot.title = element_text(face = "italic"))

p_snps_distribution_stenotrophomonas

p_snps_distribution_corynebacterium|p_snps_distribution_pseudomonas|p_snps_distribution_stenotrophomonas

```{r Fig s6a}

tree_acinetobacter <- read.tree(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\draft+ref_resolved.tre)')
rooted_tree_acinetobacter <- ape::root(tree_acinetobacter, node_number=222, resolve.root = TRUE,outgroup=222)

### display node number for each node
# ggtree(rooted_tree_acinetobacter,layout = 'circular') + geom_text(aes(subset = !isTip, label = node, x = branch), hjust = 1.2, color = "blue")

# Get the external node numbers
external_node_numbers <- 1:length(rooted_tree_acinetobacter$tip.label)
# Get the external node labels
external_node_labels <- rooted_tree_acinetobacter$tip.label
# Combine the node numbers and labels
external_node_df <- data.frame(node = external_node_numbers, label = external_node_labels)
# subset with our draft geomes node number
external_node_df_filtered <- external_node_df %>%
  filter(str_detect(label, 'coassembly'))

###color our draft genomes with their node numbers
nodes_to_color <- c(external_node_df_filtered$node)
# Group the nodes using groupOTU
tree_grouped <- groupOTU(rooted_tree_acinetobacter, nodes_to_color, "group1")

###add colored points for our draft genomes 
# Create ggtree object
ggtree_obj <- ggtree(rooted_tree_acinetobacter)
label_of_interest <- c(external_node_df_filtered$label)
subset_data <- subset(ggtree_obj$data, label %in% label_of_interest)

### circular layout
tree_circular <- ggtree(tree_grouped,layout = 'circular',aes(color=group1)) + #geom_tiplab(size=0,align = T,linetype = 'dashed',linesize = 0.1) +
  geom_tree() + geom_highlight(node=558,alpha=0.3) + #+ geom_tippoint(size=0.1)
  # geom_text2(aes(subset = !is.na(label) & isTip == FALSE, label = label), hjust = -0.5) + #show the bootstrap values
  geom_point(data = subset_data, aes(x, y), color = 'red', size = 1) +
  scale_color_manual(values = c('black','red')) + theme(legend.position = 'none')

```{r Fig s6b}

tree_klebsiella <- read.tree(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\draft+ref_resolved.tre)')
rooted_tree_klebsiella <- ape::root(tree_klebsiella, node_number=1316, resolve.root = TRUE,outgroup=1316)

ggtree(rooted_tree_klebsiella,layout = 'circular') + 
  geom_text(aes( label = node, x = branch), hjust = 1.2, color = "blue",size=1) + #show node number
  geom_text2(aes(subset = !is.na(label) & isTip == FALSE, label = label), hjust = -0.5,size=1.5) #show bootstrap values

# Get the external node numbers
external_node_numbers <- 1:length(tree_klebsiella$tip.label)
# Get the external node labels
external_node_labels <- tree_klebsiella$tip.label
# Combine the node numbers and labels
external_node_df <- data.frame(node = external_node_numbers, label = external_node_labels)
# subset with our draft geomes node number
external_node_df_filtered <- external_node_df %>%
  filter(str_detect(label, 'coassembly'))

ggtree(rooted_tree_klebsiella,layout='circular')

###color our draft genomes with their node numbers
nodes_to_color <- c(external_node_df_filtered$node)
# Group the nodes using groupOTU
tree_grouped <- groupOTU(rooted_tree_klebsiella, nodes_to_color, "group1")

###add colored points for our draft genomes 
# Create ggtree object
ggtree_obj <- ggtree(rooted_tree_klebsiella)
label_of_interest <- c(external_node_df_filtered$label)
subset_data <- subset(ggtree_obj$data, label %in% label_of_interest)

### circular layout
tree_circular <- ggtree(tree_grouped,layout = 'circular',aes(color=group1)) + #geom_tiplab(size=0,align = T,linetype = 'dashed',linesize = 0.1) +
  geom_tree() + geom_highlight(node=2102,alpha=0.3) + #+ geom_tippoint(size=0.1)
  # geom_text2(aes(subset = !is.na(label) & isTip == FALSE, label = label), hjust = -0.5) + #show the bootstrap values
  geom_point(data = subset_data, aes(x, y), color = 'red', size = 1) +
  scale_color_manual(values = c('black','red')) + theme(legend.position = 'none')

```{r Fig s6d}

### plot with Phandango

```{r Fig s6e,f}

df_recombinations_all <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\klebsiella_recombination-consensus_blocks_all.csv)')
df_pi <- read.table(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\all_patients-klebsiella-windowed.pi)",header = T)

p_recombinations <- ggplot() + 
  geom_segment(aes(x=coordinates,y=y_start,xend=coordinates,yend=y_end,size=cds_length),
               data=df_recombinations_all,color='#377eb8') +
  labs(x='Position on Chromosome',y='Recombination\nblocks') + 
  scale_x_continuous(expand = c(0.02, 0),limits=c(0,5500000),breaks=c(0,2000000,4000000),labels= scales::comma) +
  scale_y_continuous(expand = c(0, 0)) +
  # scale_y_continuous(limits = c(0,23),breaks =seq(1,23),
  #                    labels = c('P123','P143','P136','P146','P77','P25','P52', 'P112','P23',
  #                               'P26','P157','P130','P113','P118','P144','P105','P156','P141',
  #                               'P132','P114','P147','P124','P151')) + 
  # theme_pubr(base_size =30,legend = 'none',border = T) 
  theme_bw(base_size = 35) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),legend.position = 'none')

p_pi<-ggplot(df_pi,aes(x = BIN_END, y = PI))+
  geom_line(size = 0.66,color='#e41a1c') +
  theme_classic2()+
  theme(panel.spacing = unit(1.5, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        plot.margin=unit(c(1,1,1,1), "cm"))+
  scale_x_continuous(expand = c(0.02, 0),labels= scales::comma) +
  scale_y_continuous(expand = c(0, 0)) + theme_bw(base_size = 35)
p_pi <- ggpar(p_pi,
              xlab = "",
              ylab= "Pi")

```{r Fig s6g}

df_kegg_pathway = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\acinetobacter-kegg_pathway-combined.csv)")

cols <- c('Cellular Processes'='#e44235','Environmental Information Processing'='#d92aab',
          'Genetic Information Processing'='#ffc1cc','Metabolism'='#0f52ba',
          'Human Diseases'='#9baddc','Organismal Systems'='#505888')

# calculate the statistics about each category
df_kegg_pathway <- df_kegg_pathway %>% 
  count(level1,level2)
# convert gene numbers to fractions
df_kegg_pathway <- df_kegg_pathway %>%
  mutate(percentage = n/sum(n))

df_kegg_pathway <-df_kegg_pathway %>% 
  mutate(level1 = trimws(str_remove(level1, "(\\s+[A-Za-z]+)?[0-9-]+"))) %>%  
  mutate(level2 = trimws(str_remove(level2, "(\\s+[A-Za-z]+)?[0-9-]+")))

plot_kegg <-ggbarplot(df_kegg_pathway,x = "level2", y = "percentage",orientation = "horiz",
                      fill = "level1",sort.val = 'asc',sort.by.groups = F) + 
  scale_fill_manual(values=cols)
#scale_fill_manual(values=c("#e44235", "#d92aab", "#ffc1cc", "#9baddc", "#0f52ba"))

ggpar(plot_kegg,
      xlab = "",
      ylab = "Proportion",
      legend.title  = 'KEGG Class',
      legend = 'right') + theme_pubr(base_size = 24,border = F,legend = 'none')

```{r Fig s6h}

df_kegg_pathway = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\klebsiella-kegg_pathway-combined.csv)")

cols <- c('Cellular Processes'='#e44235','Environmental Information Processing'='#d92aab',
          'Genetic Information Processing'='#ffc1cc','Metabolism'='#0f52ba',
          'Human Diseases'='#9baddc','Organismal Systems'='#505888')

# calculate the statistics about each category
df_kegg_pathway <- df_kegg_pathway %>% 
  count(level1,level2)
# convert gene numbers to fractions
df_kegg_pathway <- df_kegg_pathway %>%
  mutate(percentage = n/sum(n))

df_kegg_pathway <-df_kegg_pathway %>% 
  mutate(level1 = trimws(str_remove(level1, "(\\s+[A-Za-z]+)?[0-9-]+"))) %>%  
  mutate(level2 = trimws(str_remove(level2, "(\\s+[A-Za-z]+)?[0-9-]+")))

plot_kegg <-ggbarplot(df_kegg_pathway,x = "level2", y = "percentage",orientation = "horiz",
                      fill = "level1",sort.val = 'asc',sort.by.groups = F) + 
  scale_fill_manual(values=cols)
#scale_fill_manual(values=c("#e44235", "#d92aab", "#ffc1cc", "#9baddc", "#0f52ba"))

ggpar(plot_kegg,
      xlab = "",
      ylab = "Proportion",
      legend.title  = 'KEGG Class',
      legend = 'right') + theme_pubr(base_size = 24,border = F,legend = 'none')

```{r Fig s7}

df_abundance <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\${pid}-abundance_transmitted_strains.csv)") ### replace the pid with P77,P113,P114,P130,P132 to get plot for each individual

df_abundance$Sampling_Day_ICU <- factor(df_abundance$Sampling_Day_ICU,levels = unique(df_abundance$Sampling_Day_ICU))

plot_df <- ggplot(df_abundance, aes(x=Sampling_Day_ICU, y=relative_abundance, group=taxonomy, color=taxonomy)) +
    geom_line(size=1) + 
    geom_point(size=3) +
    ggtitle(label = 'P114' ) + 
    ylim(0,100) +
    theme_bw(base_size = 14)

ggpar(plot_df,
    ylab = "Relative\nabundance (%)",
    xlab = F,
    x.text.angle = 50)  +
    theme(legend.position = "none") +
    ggeasy::easy_center_title()
